Verification of concurrent design patterns with data

### 2019-02-12 2018-Arch-pNets/arch-pNets.tex
nouveau OA
Skype meeting

For a set of variables $V \subseteq \variables$, we denote
$\valuations{V} \bydef{=} \{\val{}{}: V \rightarrow \data\}$
the set of \emph{valuations} of the variables in $V$.
Valuations extend canonically to expressions and denote $\val{}{}(e)$ the value of $e$ under the valuation $\val{}{}$.
\noteLH{its the following really necessary}
\addSB{Since we consider variables to be sorted, valuations
  applied to expressions are interpreted as partial functions
  $\exprs{V} \rightharpoonup \data$.}
For a valuation $\val{}{}$ and an assignment $(x_i := e_i)^{i \in I}$, we define:
\[
\val{}{}\bigl[(x_i := e_i)^{i \in I}\bigr](x) \bydef{=}
\begin{cases}
  \val{}{}(x), & \text{if } x \not\in x_i^{i \in I}\,,\\
  \val{}{}(e_i), & \text{if } x = x_i, \text{for some } i \in I\,.
\end{cases}
\]
For two valuations $\val{1}{}, \val{2}{} : V \rightarrow
\data$, we denote $\valdiff{\val{1}{}}{\val{2}{}} \bydef{=}
\bsetdef{x \in V}{\val{1}{}(x) \neq \val{2}{}(x)}$ the set of
variables that are assigned different values by the two
valuations.  As usual, we write $\val{1}{} \order \val{2}{}$
iff $\val{1}{}(x) \order \val{2}{}(x)$, for all $x \in V$.  An
expression $e$ is  \emph{monotonic} if, for
any two valuations $\val{1}{}, \val{2}{}$,
$\val{1}{} \order \val{2}{}$ implies $\val{1}{}(e) \order
\val{2}{}(e)$.
%
Similarly, an assignment $(x_i := e_i)^{i \in I} $ is
 monotonic if all expressions $e_i^{i \in I}$ are monotonic.
We  denote $\guards{V} \subset \boolexprs{V}$, $\monexprs{V}
\subset \exprs{V}$ and $\updates{V} \subset \assigns{V}$ the sets of
monotonic Boolean and generic expressions and assignments,
respectively.
%
%****************************************************************

@@

\noteEMin{Est-ce qu'on a la place de mettre un exemple de script SMT ici ?}

@@

%%% compositional verification
approaches by typing interfaces/protocols

infinite/parametrized model checking?

theorem proving? (too much manual effort) we need an example, we could cite Coqcots and Pycots for reconfiguration

SMT on an application not using component structure. is there a strong Z3 example

KeY? (I would call this symbolic execution and assume-guarantee)

@@

\section{pNets additional definitions}
\label{secn:pNets:additional}

\begin{definition}[Sorts, Holes, Leaves of pNets]
  \begin{itemize}
  \item  \noteSB{I think this can be lightened up a bit, since the definitions are all 
  straightforward, whereas the notation is complex.  Maybe reformulate in words and move 
  the formal definitions in the appendix?}
 The sort of a pNet is its signature, i.e. the set of actions it can
perform. \noteSB{Is there any chance that, by writing assignments explicitly ($x := e$), 
we could eliminate the need to distinguish the input variables at all? That would 
simplify things.}\hl{In the definition of sorts, we do not need to distinguish
input variables} (that specify the dataflow within LTSs), so for
computing LTS sorts, we use a substitution operator\footnote{$\subst{y_k\gets x_k}^{k\in 
K}$ is the parallel substitution 
operation.} to remove the
\emph{input marker} of variables\footnote{$\overline{S}$ denotes a set of $S$.}. Formally:
\[
\begin{array}{l}
\Sortop(\mylangle S,s_0, \to\myrangle) = \{\alpha\subst{x \gets ?x| 
x\in\symb{iv}(\alpha)}|s \xrightarrow{\langle \alpha,~g,~(x_j\!:= {e}_j)^{j\in
    J}\rangle} s'\in \to \} \\
\Sortop(\mylangle \overline{\pNet}\!, %\pNet_i^{i\in I}, \Sort_j^{j\in J}
\overline{\pNet}\!,
\overline{\symb{SV}}\myrangle)
=\{\alpha' |\, \overline{\alpha}%\alpha_j^{j\in J_k}
\to\alpha'\in\set{\symb{SV}}\}
\end{array}
\]

\item
The set of holes of a pNet is defined inductively; the sets of holes
in a pNet node and its subnets are all disjoint:
  \[\begin{array}{l}
\Holes(\mylangle S,s_0, \to\myrangle) \!=\! \emptyset \\
\Holes(\mylangle \pNet_i^{i\in I}\!,\Sort_j^{j\in J}\!, \overline{\symb{SV}}\myrangle) 
=J\uplus{\displaystyle \bigcup_{i\in 
I}\Holes(\pNet_i)}\\
\forall i\in I.\, \Holes(\pNet_i)\cap J=\emptyset\\
\forall i_1,i_2\in I.\,i_1\neq i_2\Rightarrow  
\Holes(\pNet_{i_1})\cap\Holes(\pNet_{i_2})=\emptyset
\end{array}\]
\item
The set of leaves of a pNet is the set of all pLTSs occurring in the structure, defined 
inductively as:
\[\begin{array}{l}
\Leaves(\mylangle S,s_0, \to\myrangle) \!=\!\emptyset\\%\{\mylangle S,s_0, \to\myrangle\}\\
\Leaves(\mylangle \pNet_i^{i\in I}\!,%Sort_j^{j\in J}
\overline{\Sort}\!, \overline{\symb{SV}}\myrangle) = {\displaystyle \biguplus_{i\in 
I}\Leaves(\pNet_i)\uplus\{i\mapsto \pNet_i|\pNet_i\pNet_i \text{ is a }\pLTS\}}
\end{array}\]
\end{itemize}
\end{definition}

@@

  \begin{figure}[t]
  \centering
  \includegraphics[width=\columnwidth]{TimerOAMaxProgress}
  \caption{The open Automaton of the Failure Monitor Architecture}
  \label{schema:ArchFailure:OA-MaxProgress}
\end{figure}

>>>>

For a set of variables $V \subseteq \variables$, we denote
$\valuations{V} \bydef{=} \{\val{}{}: V \rightarrow \data\}$
the set of \emph{valuations} of the variables in $V$.
Valuations extend canonically to expressions and denote $\val{}{}(e)$ the value of $e$ under the valuation $\val{}{}$.
%\noteLH{its the following really necessary}
%\addSB{Since we consider variables to be sorted, valuations
%  applied to expressions are interpreted as partial functions
%  $\exprs{V} \rightharpoonup \data$.}
For a valuation $\val{}{}$ and an assignment $(x_i := e_i)^{i \in I}$, we define:
\[
\val{}{}\bigl[(x_i := e_i)^{i \in I}\bigr](x) \bydef{=}
\begin{cases}
  \val{}{}(x), & \text{if } x \not\in x_i^{i \in I}\,,\\
  \val{}{}(e_i), & \text{if } x = x_i, \text{for some } i \in I\,.
\end{cases}
\]
For two valuations $\val{1}{}, \val{2}{} : V \rightarrow
\data$, we denote $\valdiff{\val{1}{}}{\val{2}{}} \bydef{=}
\bsetdef{x \in V}{\val{1}{}(x) \neq \val{2}{}(x)}$ the set of
variables that are assigned different values by the two
valuations.  As usual, we write $\val{1}{} \order \val{2}{}$
iff $\val{1}{}(x) \order \val{2}{}(x)$, for all $x \in V$.  An
expression $e$ is  \emph{monotonic} if, for
any two valuations $\val{1}{}, \val{2}{}$,
$\val{1}{} \order \val{2}{}$ implies $\val{1}{}(e) \order
\val{2}{}(e)$.
%
Similarly, an assignment $(x_i := e_i)^{i \in I} $ is
 monotonic if all expressions $e_i^{i \in I}$ are monotonic.
We  denote $\guards{V} \subset \boolexprs{V}$, $\monexprs{V}
\subset \exprs{V}$ and $\updates{V} \subset \assigns{V}$ the sets of
monotonic Boolean and generic expressions and assignments,
respectively.
%
%****************************************************************

@@

%\noteEMin{Est-ce qu'on a la place de mettre un exemple de script SMT ici ?}

@@

%%% compositional verification
approaches by typing interfaces/protocols == Simon

infinite/parametrized model checking? == ERIC + Simon

theorem proving? (too much manual effort) we need an example, we could cite Coqcots and Pycots for reconfiguration == LUDO

SMT on an application not using component structure. is there a strong Z3 example == Eric

KeY? (I would call this symbolic execution and assume-guarantee) == LUDO

@@

%*********************************************************************
%*********************************************************************
%
%\section{pNets additional definitions}
%\label{secn:pNets:additional}
%
%\begin{definition}[Sorts, Holes, Leaves of pNets]
%  \begin{itemize}
%  \item  \noteSB{I think this can be lightened up a bit, since the definitions are all 
%  straightforward, whereas the notation is complex.  Maybe reformulate in words and move 
%  the formal definitions in the appendix?}
% The sort of a pNet is its signature, i.e. the set of actions it can
%perform. \noteSB{Is there any chance that, by writing assignments explicitly ($x := e$), 
%we could eliminate the need to distinguish the input variables at all? That would 
%simplify things.}\hl{In the definition of sorts, we do not need to distinguish
%input variables} (that specify the dataflow within LTSs), so for
%computing LTS sorts, we use a substitution operator\footnote{$\subst{y_k\gets x_k}^{k\in 
%K}$ is the parallel substitution 
%operation.} to remove the
%\emph{input marker} of variables\footnote{$\overline{S}$ denotes a set of $S$.}. Formally:
%\[
%\begin{array}{l}
%\Sortop(\mylangle S,s_0, \to\myrangle) = \{\alpha\subst{x \gets ?x| 
%x\in\symb{iv}(\alpha)}|s \xrightarrow{\langle \alpha,~g,~(x_j\!:= {e}_j)^{j\in
%    J}\rangle} s'\in \to \} \\
%\Sortop(\mylangle \overline{\pNet}\!, %\pNet_i^{i\in I}, \Sort_j^{j\in J}
%\overline{\pNet}\!,
%\overline{\symb{SV}}\myrangle)
%=\{\alpha' |\, \overline{\alpha}%\alpha_j^{j\in J_k}
%\to\alpha'\in\set{\symb{SV}}\}
%\end{array}
%\]
%
%\item
%The set of holes of a pNet is defined inductively; the sets of holes
%in a pNet node and its subnets are all disjoint:
%  \[\begin{array}{l}
%\Holes(\mylangle S,s_0, \to\myrangle) \!=\! \emptyset \\
%\Holes(\mylangle \pNet_i^{i\in I}\!,\Sort_j^{j\in J}\!, \overline{\symb{SV}}\myrangle) 
%=J\uplus{\displaystyle \bigcup_{i\in 
%I}\Holes(\pNet_i)}\\
%\forall i\in I.\, \Holes(\pNet_i)\cap J=\emptyset\\
%\forall i_1,i_2\in I.\,i_1\neq i_2\Rightarrow  
%\Holes(\pNet_{i_1})\cap\Holes(\pNet_{i_2})=\emptyset
%\end{array}\]
%\item
%The set of leaves of a pNet is the set of all pLTSs occurring in the structure, defined 
%inductively as:
%\[\begin{array}{l}
%\Leaves(\mylangle S,s_0, \to\myrangle) \!=\!\emptyset\\%\{\mylangle S,s_0, \to\myrangle\}\\
%\Leaves(\mylangle \pNet_i^{i\in I}\!,%Sort_j^{j\in J}
%\overline{\Sort}\!, \overline{\symb{SV}}\myrangle) = {\displaystyle \biguplus_{i\in 
%I}\Leaves(\pNet_i)\uplus\{i\mapsto \pNet_i|\pNet_i\pNet_i \text{ is a }\pLTS\}}
%\end{array}\]
%\end{itemize}
%\end{definition}

@@

  \begin{figure}[t]
  \centering
  \includegraphics[width=\columnwidth]{TimerOAMaxProgress}
  \caption{The open Automaton of the Failure Monitor Architecture,
    with Maximal Progress}
  \label{schema:ArchFailure:OA-MaxProgress}
\end{figure}
